### 2021-11-18 R0-Density-Reanalysis/R0-Density-Reanalysis.Rmd
Reorganized the repository after paper's publication online at Geographical Analysis

---
title: "Reproducibility of research during COVID-19: examining the case of population density and the basic reproductive rate from the perspective of spatial analysis"
author:
  - name: Author
    email: author@institution.edu
    affiliation: University
    footnote: 1
address:
  - code: University
    address: Department, Street, City, State ZIP
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
  The emergence of the novel SARS-CoV-2 coronavirus and the global COVID-19 pandemic in 2019 led to explosive growth in scientific research. Alas, much of the research in the literature lacks conditions to be reproducible, and recent publications on the association between population density and the basic reproductive number of SARS-CoV-2 are no exception. Relatively few papers share code and data sufficiently, which hinders not only verification but additional experimentation. In this paper, an example of reproducible research shows the potential of spatial analysis for epidemiology research during COVID-19. Transparency and openness means that independent researchers can, with only modest efforts, verify findings and use different approaches as appropriate. Given the high stakes of the situation, it is essential that scientific findings, on which good policy depends, are as robust as possible; as the empirical example shows, reproducibility is one of the keys to ensure this.

journal: "Geographical Analysis"
date: "`r Sys.Date()`"
bibliography: [bibliography.bib, packages.bib]
linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
header-includes:
    - \usepackage{setspace}\onehalfspacing
---

>>>>

---
title: "Reproducibility of research during COVID-19: examining the case of population density and the basic reproductive rate from the perspective of spatial analysis"
author:
  - name: Antonio Paez
    email: paezha@mcmaster.ca
    affiliation: McMaster University
    footnote: 1
address:
  - code: McMaster University
    address: School of Earth, Environment and Society, 1280 Main St West, Hamilton, Ontario L8S 4K1 Canada
footnote:
  - code: 1
    text: "Corresponding Author"
abstract: |
  The emergence of the novel SARS-CoV-2 coronavirus and the global COVID-19 pandemic in 2019 led to explosive growth in scientific research. Alas, much of the research in the literature lacks conditions to be reproducible, and recent publications on the association between population density and the basic reproductive number of SARS-CoV-2 are no exception. Relatively few papers share code and data sufficiently, which hinders not only verification but additional experimentation. In this paper, an example of reproducible research shows the potential of spatial analysis for epidemiology research during COVID-19. Transparency and openness means that independent researchers can, with only modest efforts, verify findings and use different approaches as appropriate. Given the high stakes of the situation, it is essential that scientific findings, on which good policy depends, are as robust as possible; as the empirical example shows, reproducibility is one of the keys to ensure this.

  This paper is now published in Geographical Analysis (\url{https://doi.org/10.1111/gean.12307})

journal: "Geographical Analysis"
date: "`r Sys.Date()`"
bibliography: bibliography.bib
linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
header-includes:
    - \usepackage{setspace}\onehalfspacing    
---

### 2021-11-18 R0-Density-Reanalysis/R0-Density-Reanalysis.tex
Reorganized the repository after paper's publication online at Geographical Analysis

  \title{Reproducibility of research during COVID-19: examining the case
of population density and the basic reproductive rate from the
perspective of spatial analysis}
    \author[University]{Author\corref{1}}
   \ead{author@institution.edu} 
      \address[University]{Department, Street, City, State ZIP}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  The emergence of the novel SARS-CoV-2 coronavirus and the global
  COVID-19 pandemic in 2019 led to explosive growth in scientific
  research. Alas, much of the research in the literature lacks
  conditions to be reproducible, and recent publications on the
  association between population density and the basic reproductive
  number of SARS-CoV-2 are no exception. Relatively few papers share
  code and data sufficiently, which hinders not only verification but
  additional experimentation. In this paper, an example of reproducible
  research shows the potential of spatial analysis for epidemiology
  research during COVID-19. Transparency and openness means that
  independent researchers can, with only modest efforts, verify findings
  and use different approaches as appropriate. Given the high stakes of
  the situation, it is essential that scientific findings, on which good
  policy depends, are as robust as possible; as the empirical example
  shows, reproducibility is one of the keys to ensure this.
  \end{abstract}

@@

\caption{\label{tab:tabulate-swn-results}\label{tab:swn-results}Reproducing SWN: Models 1-3}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllllrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Model 1} & \multicolumn{2}{c}{Model 2} & \multicolumn{2}{c}{Model 3} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variable & beta & 95\% CI & beta & 95\% CI & beta & 95\% CI\\
\midrule
Intercept & 2.274 & [2.167, 2.381] & 3.347 & [2.676, 4.018] & 3.386 & [2.614, 4.157]\\
Log of population density & 0.162 & [0.133, 0.191] & 0.145 & [0.115, 0.176] & 0.147 & [0.113, 0.18]\\
Percent of private transportation &  &  & -0.013 & [-0.02, -0.005] & -0.013 & [-0.021, -0.005]\\
Median household income (\$10,000) &  &  &  &  & -0.003 & [-0.033, 0.026]\\
Standard deviation (Intercept) & 0.166 & [0.108, 0.254] & 0.136 & [0.081, 0.229] & 0.137 & [0.081, 0.232]\\
\addlinespace
Within-group standard error & 0.665 & [0.638, 0.693] & 0.665 & [0.638, 0.693] & 0.665 & [0.638, 0.694]\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:tabulate-sample-selection-results}\label{tab:selection-results}Estimation results of sample selection models}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Selection Model 1} & \multicolumn{2}{c}{Selection Model 2} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & $\beta$ & 95\% CI & $\beta$ & 95\% CI\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Sample Selection Model}}\\
\hspace{1em}Intercept & -2.237 & [-3.109, -1.365] & -7.339 & [-8.381, -6.297]\\
\hspace{1em}Log of population density & 0.385 & [0.352, 0.418] &  & \\
\hspace{1em}Density (1,000 per sq.km) &  &  & 2.484 & [2.13, 2.838]\\
\hspace{1em}Density squared &  &  & -0.387 & [-0.473, -0.3]\\
\hspace{1em}Percent of private transportation & 0.025 & [0.016, 0.034] & 0.057 & [0.046, 0.067]\\
\hspace{1em}Median household income (10,000) & 0.202 & [0.168, 0.235] & 0.32 & [0.283, 0.357]\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Outcome Model}}\\
\hspace{1em}Intercept & 0.605 & [-0.257, 1.466] & 2.784 & [1.652, 3.915]\\
\hspace{1em}Log of population density & 0.39 & [0.354, 0.426] &  & \\
\hspace{1em}Density (1,000 per sq.km) &  &  & 0.758 & [0.509, 1.008]\\
\hspace{1em}Density squared &  &  & -0.132 & [-0.187, -0.077]\\
\hspace{1em}Percent of private transportation & 0.01 & [0.001, 0.018] & -0.011 & [-0.021, -0.001]\\
\hspace{1em}Median household income (\$10,000) & 0.126 & [0.094, 0.159] & 0.002 & [-0.033, 0.037]\\
$\sigma$ & 0.954 & [0.904, 1.003] & 0.684 & [0.652, 0.716]\\
$\rho$ & 0.971 & [0.961, 0.98] & -0.199 & [-0.377, -0.022]\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:tabulate-sample-selection3-results}\label{tab:selection3-results}Estimation results of sample selection model with spatial filter}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Selection Model 3} \\
\cmidrule(l{3pt}r{3pt}){2-3}
Variable & $\beta$ & 95\% CI\\
\midrule
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Sample Selection Model}}\\
\hspace{1em}Intercept & -7.304 & [-8.346, -6.262]\\
\hspace{1em}Density (1,000 per sq.km) & 2.445 & [2.089, 2.802]\\
\hspace{1em}Density squared & -0.380 & [-0.468, -0.292]\\
\hspace{1em}Percent of private transportation & 0.056 & [0.046, 0.067]\\
\hspace{1em}Median household income (10,000) & 0.318 & [0.28, 0.356]\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Outcome Model}}\\
\hspace{1em}Intercept & 2.563 & [2.497, 2.629]\\
\hspace{1em}Density (1,000 per sq.km) & 0.760 & [0.746, 0.774]\\
\hspace{1em}Density squared & -0.133 & [-0.135, -0.13]\\
\hspace{1em}Percent of private transportation & -0.011 & [-0.012, -0.011]\\
\hspace{1em}Median household income (\$10,000) & 0.002 & [-0.001, 0.004]\\
\hspace{1em}Spatial filter & 1.000 & [0.998, 1.001]\\
$\sigma$ & 0.017 & [0.015, 0.019]\\
$\rho$ & -0.304 & [-0.957, 0.349]\\
\bottomrule
\end{tabular}
\end{table}

@@

>>>>

  \title{Reproducibility of research during COVID-19: examining the case
of population density and the basic reproductive rate from the
perspective of spatial analysis}
    \author[McMaster University]{Antonio Paez\corref{1}}
   \ead{paezha@mcmaster.ca} 
      \address[McMaster University]{School of Earth, Environment and
Society, 1280 Main St West, Hamilton, Ontario L8S 4K1 Canada}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  The emergence of the novel SARS-CoV-2 coronavirus and the global
  COVID-19 pandemic in 2019 led to explosive growth in scientific
  research. Alas, much of the research in the literature lacks
  conditions to be reproducible, and recent publications on the
  association between population density and the basic reproductive
  number of SARS-CoV-2 are no exception. Relatively few papers share
  code and data sufficiently, which hinders not only verification but
  additional experimentation. In this paper, an example of reproducible
  research shows the potential of spatial analysis for epidemiology
  research during COVID-19. Transparency and openness means that
  independent researchers can, with only modest efforts, verify findings
  and use different approaches as appropriate. Given the high stakes of
  the situation, it is essential that scientific findings, on which good
  policy depends, are as robust as possible; as the empirical example
  shows, reproducibility is one of the keys to ensure this.

  This paper is now published in Geographical Analysis
  (\url{https://doi.org/10.1111/gean.12307})
  \end{abstract}

@@

\caption{\label{tab:tabulate-swn-results}\label{tab:swn-results}Reproducing SWN: Models 1-3}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllllrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Model 1} & \multicolumn{2}{c}{Model 2} & \multicolumn{2}{c}{Model 3} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variable & beta & 95\% CI & beta & 95\% CI & beta & 95\% CI\\
\midrule
Intercept & 2.274 & {}[2.167, 2.381] & 3.347 & {}[2.676, 4.018] & 3.386 & {}[2.614, 4.157]\\
Log of population density & 0.162 & {}[0.133, 0.191] & 0.145 & {}[0.115, 0.176] & 0.147 & {}[0.113, 0.18]\\
Percent of private transportation &  &  & -0.013 & {}[-0.02, -0.005] & -0.013 & {}[-0.021, -0.005]\\
Median household income (\$10,000) &  &  &  &  & -0.003 & {}[-0.033, 0.026]\\
Standard deviation (Intercept) & 0.166 & {}[0.108, 0.254] & 0.136 & {}[0.081, 0.229] & 0.137 & {}[0.081, 0.232]\\
\addlinespace
Within-group standard error & 0.665 & {}[0.638, 0.693] & 0.665 & {}[0.638, 0.693] & 0.665 & {}[0.638, 0.694]\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:tabulate-sample-selection-results}\label{tab:selection-results}Estimation results of sample selection models}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Selection Model 1} & \multicolumn{2}{c}{Selection Model 2} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & $\beta$ & 95\% CI & $\beta$ & 95\% CI\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Sample Selection Model}}\\
\hspace{1em}Intercept & -2.237 & {}[-3.109, -1.365] & -7.339 & {}[-8.381, -6.297]\\
\hspace{1em}Log of population density & 0.385 & {}[0.352, 0.418] &  & \\
\hspace{1em}Density (1,000 per sq.km) &  &  & 2.484 & {}[2.13, 2.838]\\
\hspace{1em}Density squared &  &  & -0.387 & {}[-0.473, -0.3]\\
\hspace{1em}Percent of private transportation & 0.025 & {}[0.016, 0.034] & 0.057 & {}[0.046, 0.067]\\
\hspace{1em}Median household income (10,000) & 0.202 & {}[0.168, 0.235] & 0.32 & {}[0.283, 0.357]\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Outcome Model}}\\
\hspace{1em}Intercept & 0.605 & {}[-0.257, 1.466] & 2.784 & {}[1.652, 3.915]\\
\hspace{1em}Log of population density & 0.39 & {}[0.354, 0.426] &  & \\
\hspace{1em}Density (1,000 per sq.km) &  &  & 0.758 & {}[0.509, 1.008]\\
\hspace{1em}Density squared &  &  & -0.132 & {}[-0.187, -0.077]\\
\hspace{1em}Percent of private transportation & 0.01 & {}[0.001, 0.018] & -0.011 & {}[-0.021, -0.001]\\
\hspace{1em}Median household income (\$10,000) & 0.126 & {}[0.094, 0.159] & 0.002 & {}[-0.033, 0.037]\\
$\sigma$ & 0.954 & {}[0.904, 1.003] & 0.684 & {}[0.652, 0.716]\\
$\rho$ & 0.971 & {}[0.961, 0.98] & -0.199 & {}[-0.377, -0.022]\\
\bottomrule
\end{tabular}}
\end{table}

@@

\caption{\label{tab:tabulate-sample-selection3-results}\label{tab:selection3-results}Estimation results of sample selection model with spatial filter}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Selection Model 3} \\
\cmidrule(l{3pt}r{3pt}){2-3}
Variable & $\beta$ & 95\% CI\\
\midrule
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Sample Selection Model}}\\
\hspace{1em}Intercept & -7.249 & {}[-8.285, -6.214]\\
\hspace{1em}Density (1,000 per sq.km) & 2.424 & {}[2.074, 2.774]\\
\hspace{1em}Density squared & -0.373 & {}[-0.459, -0.288]\\
\hspace{1em}Percent of private transportation & 0.056 & {}[0.045, 0.066]\\
\hspace{1em}Median household income (10,000) & 0.319 & {}[0.282, 0.356]\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Outcome Model}}\\
\hspace{1em}Intercept & 2.290 & {}[2.026, 2.553]\\
\hspace{1em}Density (1,000 per sq.km) & 0.843 & {}[0.786, 0.9]\\
\hspace{1em}Density squared & -0.142 & {}[-0.153, -0.131]\\
\hspace{1em}Percent of private transportation & -0.010 & {}[-0.012, -0.008]\\
\hspace{1em}Median household income (\$10,000) & 0.011 & {}[0.003, 0.02]\\
\hspace{1em}Spatial filter & 1.001 & {}[0.992, 1.011]\\
$\sigma$ & 0.120 & {}[0.107, 0.133]\\
$\rho$ & 0.495 & {}[0.218, 0.772]\\
\bottomrule
\end{tabular}
\end{table}

@@

\hypertarget{acknwledgments}{%
\section*{Acknowledgments}\label{acknwledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

The analysis reported in this paper was conducted in the \texttt{R}
computing statistical language (\textbf{R-base?}). The source document
is an Rmarkdown document (\textbf{rmarkdown2018?};
\textbf{rmarkdown2020?}) processed using \texttt{knitr}
(\textbf{knitr2015?}; \textbf{knitr2014?}). The following packages were
used in the analysis, and I wish to acknowledge their creators for their
generous efforts: \texttt{adespatial} (\textbf{R-adespatial?}),
\texttt{censReg} (\textbf{R-censReg?}), \texttt{dplyr}
(\textbf{R-dplyr?}), \texttt{forcats} (\textbf{R-forcats?}),
\texttt{ggplot2} (\textbf{ggplot22016?}), \texttt{gmm}
(\textbf{gmm2010?}), \texttt{kableExtra} (\textbf{R-kableExtra?}),
\texttt{Matrix} (\textbf{R-Matrix?}), \texttt{maxLik}
(\textbf{maxLik2011?}), \texttt{miscTools} (\textbf{R-miscTools?}),
\texttt{mvtnorm} (\textbf{mvtnorm2009?}), \texttt{nlme}
(\textbf{R-nlme?}), \texttt{patchwork} (\textbf{R-patchwork?}),
\texttt{purrr} (\textbf{R-purrr?}), \texttt{readr} (\textbf{R-readr?}),
\texttt{sampleSelection} (\textbf{sampleSelection2008?}),
\texttt{sandwich} (\textbf{sandwich2020?}; \textbf{sandwich2004?};
\textbf{sandwich2006?}), \texttt{scico} (\textbf{R-scico?}), \texttt{sf}
(\textbf{sf2018?}), \texttt{sp} (\textbf{sp2005?}; \textbf{sp2013?}),
\texttt{spatialprobit} (\textbf{R-spatialprobit?}), \texttt{spData}
(\textbf{R-spData?}), \texttt{spdep} (\textbf{spdep2018?};
\textbf{spdep2013?}), \texttt{stringr} {[}R-stringr{]}, \texttt{tibble}
(\textbf{R-tibble?}), \texttt{tidycensus} (\textbf{R-tidycensus?}),
\texttt{tidyr} (\textbf{R-tidyr?}), \texttt{tidyverse}
(\textbf{tidyverse2019?}), \texttt{tmvtnorm} (\textbf{R-tmvtnorm?}),
\texttt{units} (\textbf{units2016?}). This research was not supported by
Canada's Research Councils.

### 2021-11-18 README.Rmd
Reorganized the repository after paper's publication online at Geographical Analysis

Paper desk-rejected by [~~PLoS ONE~~](https://github.com/paezha/Reproductive-Rate-and-Density-US-Reanalyzed/tree/main/R0-Density-Reanalysis/PLOS ONE Decision_ PONE-D-21-16874.pdf) [~~Science of the Total Environment~~](https://github.com/paezha/Reproductive-Rate-and-Density-US-Reanalyzed/tree/main/R0-Density-Reanalysis/Your Submission STOTEN-D-21-12077.pdf) [~~Landscape and Urban Planning~~](https://github.com/paezha/Reproductive-Rate-and-Density-US-Reanalyzed/tree/main/R0-Density-Reanalysis/Decision on submission to Landscape and Urban Planning.pdf)

Forthcoming in Geographical Analysis

>>>>

Paper desk-rejected by [~~PLoS ONE~~](https://github.com/paezha/Reproductive-Rate-and-Density-US-Reanalyzed/tree/main/R0-Density-Reanalysis/Editorial-Materials-and-Correspondence/PLOS ONE Decision_ PONE-D-21-16874.pdf) [~~Science of the Total Environment~~](https://github.com/paezha/Reproductive-Rate-and-Density-US-Reanalyzed/tree/main/R0-Density-Reanalysis/Editorial-Materials-and-Correspondence/Your Submission STOTEN-D-21-12077.pdf) [~~Landscape and Urban Planning~~](https://github.com/paezha/Reproductive-Rate-and-Density-US-Reanalyzed/tree/main/R0-Density-Reanalysis/Editorial-Materials-and-Correspondence/Decision on submission to Landscape and Urban Planning.pdf)

Geographical Analysis (2021) https://doi.org/10.1111/gean.12307
